Using Metric Space Indexing for Complete and Efficient Record Linkage

### 2017-11-22 paper.tex
Update on Overleaf.

\begin{figure}[!t]
  \centering
  \includegraphics[width=1.0\textwidth]{figures/linkage-process}
  \caption{Overview of the steps of the traditional record linkage
           process (left-hand side) and our proposed metric space
           indexing based approach (right-hand side), as described in
           Sect.~\ref{sec-intro}, where records from two datasets,
           $\mathbf{D}_A$ and $\mathbf{D}_B$, are being linked.}
           \label{fig-rl-process}
\end{figure}

@@

Metric space indexing (MSI) techniques provide indexing structures to
support the comparison of records in one set with those in another,
typically also offering similarity search operations. To create an MSI
data structure, it is necessary to define a distance measure between
records, with certain properties including the \emph{triangle
inequality}. Similarity search operations include
\emph{range-search(\textbf{q},d)}, where all records $\mathbf{r}$
in a database $\mathbf{D}$ within a distance $d$ of a query record
$\mathbf{q}$ are identified; \emph{nearest-neighbour(\textbf{q})},
returning the record $\mathbf{r} \in \mathbf{D}$ with smallest
distance to $\q$; and \emph{nearest-n(\textbf{q},n)}, returning
the $n$ closest records in $\mathbf{D}$ to $\mathbf{q}$.
\cite{Zezula2010} gives an overview of MSI techniques. Here we choose
one metric space indexing structure, the M-tree~\cite{paolociaccia2m},
and investigate its efficacy for record linkage.

@@

We address the following general record linkage problem: for two
datasets $A$ and $B$, we wish to find, for each record in $A$, all the
records in $B$ that match it. We compare several algorithms for record
linkage: traditional blocking, one incomplete similarity search method,
LSH-MinHash, and one complete method, M-tree~\cite{paolociaccia2m}. We
also use a simple complete brute force technique as a baseline for
comparison, although this can only feasibly be applied to the smallest
of our datasets. All experiments have a number of parameters to
configure the search space and algorithm behaviour, including a
\emph{distance threshold} specifying a maximum distance, equivalent to a
minimum similarity, for two records to be classified as a link (i.e.\
referring to the same entity).

\textbf{Brute force:}
Two nested loops are used to compare every record in $A$ with every
record in $B$. Each pair is classified as a link if the distance between
the records is less than the threshold. This approach is guaranteed to
identify all links, with complexity $ O(|A| \times |B|) $.

\textbf{Traditional blocking:}
The additional parameters are the set of blocking keys and the
normalization encodings applied to each field. These are selected as
described in \cite{Chr12b}, exploiting knowledge of the domain and of
the records being linked, and chosen with the intention of giving the
best possible results for the datasets. Each record in $A$ is placed
into the appropriate block based on its blocking key value. The
algorithm then iterates over the records in $B$, and for each one,
compares it with each of the records from $A$ in the block with the same
blocking key value.

\textbf{LSH-MinHash:}
The additional parameters are \emph{shingle size} ($l_{ss}$), \emph{band
size} ($l_{bs}$) and \emph{number of bands} ($l_{nb}$). Each record in
$A$ is placed in an \emph{LSH-MinHash} data structure. First, all the
fields of the record are concatenated, and the result \emph{shingled}
into a set of n-grams with $n = l_{ss}$. Next, a set of
deterministically generated hash functions are applied to each n-gram in
the set and the smallest result (the MinHash) of each hash application
is added to a signature for the record. The number of hashes used, and
thus the size of the signature, is set to $l_{nb} \times l_{bs}$.
Finally, the signature is split into $l_{nb}$ bands and the values from
each band are hashed again to create a number of keys. The original
record is added to a map associated with each of the keys.

To perform linkage, the algorithm iterates over the records in $B$. Each
record is hashed using the procedure described above, to obtain a set of
keys. For each of these keys, the key is looked up in the LSH-MinHash
data structure, and the records from $A$ associated with it added to the
result set. Finally, the record from $B$ is compared in turn with each
record in the result set, with the pair being classified as a link or
non-link based on their distance.

In some circumstances, incomplete approaches such as traditional
blocking and LSH-MinHash can yield higher precision than complete
techniques. This can occur when the datasets contain a significant
number of non-matches that nonetheless have high similarity. In this
situation, the fact that an incomplete technique omits consideration of
some potential links can serve to improve precision, since a
classification decision based on similarity threshold is incorrect for
high-similarity non-matches. By definition, recall is never higher for
incomplete techniques.

\textbf{M-tree:} The M-tree linkage algorithm has no additional
parameters. In a similar manner to the \emph{LSH-MinHash} approach, each
record in A is inserted into an M-tree. To perform linkage, the
algorithm iterates over each record $b$ in $B$. A \emph{range-search(b,d)}
operation is performed, passing the distance threshold as the second
parameter. All the returned records are directly recorded as links.

@@

\begin{table}[t]
\caption{Characteristics of datasets used in the experiments.}
 \label{table-datasets}
  \centering
  \begin{scriptsize}
  %\addtolength{\tabcolsep}{-0.5pt}
  \begin{tabular}{ccccc}
  \hline\noalign{\smallskip}
  Dataset~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities\\
  name(s)  & dataset 1  & dataset 2  & matching pairs & linked \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
  Cora & 1,295 & 1,295 & 17,184 & publication-publication\\
  Isle of Skye & 17,612 & 12,284& 2,900 & birth-death\\
  Kilmarnock & 38,430 & 23,714 & 8,300 & birth-death\\
  \noalign{\smallskip} \hline
  \end{tabular}
  \end{scriptsize}
\end{table}

@@

\smallskip
\textbf{Datasets:}
\label{sec-data}
We used three datasets from two domains in our experiments, as
summarised in Table~\ref{table-datasets}. The first is
\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer to
112 machine learning publications. It is used as a benchmark dataset in
the literature for assessing linkage algorithms. Ground truth is
provided via a unique \emph{paper\_id} identifier of the form
``blum1993". In this experiment linkage is performed over the same set
of records.

The other datasets are historical Scottish records of vital events
(birth, marriages and deaths) one registered on the \emph{Isle of Skye},
a rural district, and the other records from \emph{Kilmarnock}, an
industrial town. These datasets were create by historical demographers
who extensively curated and linked both
datasets~\cite{reid2002,reid2006}. Both datasets include the names,
gender, addresses of individuals and their parents. Ground truth was
generated by the demographers based on their extensive domain knowledge.

In all of our experiments we use a single distance metric which is the
sum of the Levenshtein~\cite{Levenshtein66} distances between the set of
fields being compared.

@@

In this section, we perform linkage on the Cora dataset using all four
approaches presented in this paper: brute force, traditional blocking,
LSH and M-tree, using several selected configurations for the blocking
and LSH linkers. The distance threshold is varied between 0 and
250~\footnote{Relatively high Levenshtein distances are included since
Cora includes a number of low-similarity true matches.}. For traditional
blocking, the following Cora fields are used individually as blocking
keys: \textit{author}, \textit{title}, \textit{venue},
\textit{location}, \textit{publisher} and \textit{year}. We also use a
combined blocking key comprising all fields.

\Cref{cora-quality} shows the precision, recall, and F-measure for
various thresholds. As expected, low thresholds give high precision and
low recall, and the reverse for high threshold values. Brute force and
M-tree give identical results. The best linkage quality, with an
F-measure of around 0.7, is achieved by a number of linkers, including
brute force, M-tree, blocking on `authors' and blocking and on all
fields, and two of the LSH configurations. All of these give similar
overall results, apart from blocking on `authors', which gives much
better quality at very high distance threshold values. This is due to
the incomplete nature of the linker, avoiding consideration of
significant numbers of high-similarity non-matches, thus avoiding these
becoming false positives and keeping precision high.

For a more detailed investigation of selected linkers, the brute force
approach is used to establish a good threshold value for the Cora
dataset. The maximum F-measure is observed at a threshold value of
70. This value is dataset-dependent; for different datasets the maximum
F-measure will occur at different thresholds. In the rest of this
section we fix the threshold value at 70.

\Cref{comparison-of-results-cora} shows greater detail for selected
linkers, showing the parameters for the experiment, the number of
distance comparisons made, the precision, recall and the F-measure
achieved by each algorithm.  In the \emph{Linker} column the algorithm
name is followed by its parameters: for LSH the number of the bands
followed by the band size, and for traditional blocking the fields used
for blocking. The number of distance comparisons is reported as a
machine-independent proxy for execution cost, since code profiling shows
that distance calculations are dominant.

@@

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68       \\ \hline
M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\ \hline
LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
LSH-5-2            &  342,849	& 0.91	& 0.55	& 0.69       \\
LSH-10-2           &  513,947	& 0.88	& 0.57	& 0.69       \\
LSH-2-5            &  14,329	& 0.99	& 0.28	& 0.43      \\
LSH-5-5            &  22,057	& 0.99	& 0.36	& 0.53       \\
LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
LSH-2-10           &  4,711	& 1	& 0.15	& 0.27    \\
LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\  \hline
Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
Block-authors      &   11,039                   & 0.94      & 0.16   & 0.28       \\
Block-title        &   27,407                   & 0.95      & 0.42   & 0.58       \\
Block-venue        &   36,647                   & 0.85      & 0.29   & 0.44       \\
Block-location     & 1,009,957                 & 0.83      & 0.43   & 0.57       \\
Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\ \hline
\end{tabular}
\caption{Linkage quality on Cora dataset with distance threshold = 70
\label{comparison-of-results-cora}
}
\end{table}

@@

GRAHAM STILL WRITING HERE

Birth records were linked to death records, separately for the Skye and Kilmarnock datasets, using M-tree and LSH.

It was not computationally feasible to run the brute force linker. Of the incomplete linkers, LSH was selected as it gave slightly better results for Cora.

The shingle size was set to 2 for all the LSH experiments reported, as this was found to give good results and LSH was not especially sensitive to this parameter. Results for other shingle sizes are omitted from this paper for brevity.

on the Cora dataset using all four
approaches presented in this paper: brute force, traditional blocking,
LSH and M-tree, using several selected configurations for the blocking
and LSH linkers. The distance threshold is varied between 0 and
250~\footnote{Relatively high Levenshtein distances are included since
Cora includes a number of low-similarity true matches.}. For traditional
blocking, the following Cora fields are used individually as blocking
keys: \textit{author}, \textit{title}, \textit{venue},
\textit{location}, \textit{publisher} and \textit{year}. We also use a
combined blocking key comprising all fields.

\Cref{cora-quality} shows the precision, recall, and F-measure for
various thresholds. As expected, low thresholds give high precision and
low recall, and the reverse for high threshold values. Brute force and
M-tree give identical results. The best linkage quality, with an
F-measure of around 0.7, is achieved by a number of linkers, including
brute force, M-tree, blocking on `authors' and blocking and on all
fields, and two of the LSH configurations. All of these give similar
overall results, apart from blocking on `authors', which gives much
better quality at very high distance threshold values. This is due to
the incomplete nature of the linker, avoiding consideration of
significant numbers of high-similarity non-matches, thus avoiding these
becoming false positives and keeping precision high.

\Cref{demography-quality}

\Cref{comparison-of-results-demography-skye} and \Cref{comparison-of-results-demography-kili} show the results for LSH-MinHash and M-tree on the Skye and Kilmarnock datasets.

In both cases the F1 measure achieved is better for M-tree than any of the experimental settings chosen for LSH. This is largely due to the recall for M-tree being much higher than any of the LSH experiments. For most configurations, LSH out-performs M-tree in terms of precision.

Perhaps more important than the fact that M-tree gives better F1 measures than all of the LSH experiments is the fact that the LSH performance is extremely variable. For a relatively plausible settings for the number of bands and band-size the F1 measurements vary from 0.01 (extremely poor) to 0.47 (relatively good) for Skye and between 0.03 and 0.49 for the Kilmarnock dataset. In both cases LSH-10-2 performs best, but since this is data dependent there is no guarantee these parameters would work well with another dataset. Furthermore, without good ground truth, which is usually unavailable, there is no way of knowing if good linkage results are being obtained or not.

Another observation that is worthy of comment is the number of (distance) comparisons being performed by the different algorithms. For M-trees the number of comparisons are closely related to the complexity of the search being performed since the algorithms are performing (range) searches over distance measures to determine the query results. This is not true of LSH which is performing Jaccard similarity using Hashes over the datasets. Comparisons are only used at the final step to determine if a candidate is a match or not. The LSH results which provide the best results by F1 measure are performing between a quarter and half the number of matches performed by M-trees. From this we can conclude that the hash algorithms are returning many candidates that are outwith the search distance. Thus, in order to get good results, LSH is tending towards a brute force search over the candidate results. Despite this LSH is much faster due to the efficiency of the hashing mechanisms.

@@

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
\caption{Linkage quality on Skye dataset using M-tree\label{skye-quality-mtree}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
\caption{Linkage quality on Kilmarnock dataset using M-tree\label{kilmarnock-quality-mtree}}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
\caption{Linkage quality on Skye dataset using M-tree and all LSH configurations}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
\caption{Linkage quality on Kilmarnock dataset using M-tree and all LSH configurations}
\end{subfigure}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\ \hline

LSH-2-2  & 3,109,250                   & 0.63      & 0.03   & 0.06       \\

@@

LSH-10-5 & 129,873                    & 0.72      & 0.01   & 0.02       \\ \hline
\end{tabular}
\caption{Linkage quality on Skye dataset with distance threshold = 2
\label{comparison-of-results-demography-skye}
}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\ \hline
LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338                   & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848                   & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ \hline
\end{tabular}
\caption{Linkage quality on Kilmarnock dataset with distance threshold = 2
\label{comparison-of-results-demography-kili}
}
\end{table}

@@

In this paper we have demonstrated the efficacy of MSI in achieving  complete and efficient record linkage, without the need for complex parameter tuning. In conclusion, this assertion deserves some careful unpacking. It is always possible to achieve high quality linkage using a brute force approach. However the n-squared complexity of this approach prevents its practical application for datasets of even moderate size. The use of MSI techniques such as M-tree delivers the same results as brute force with fewer comparisons. We have shown that the M-tree approach delivers high precision, high recall results that are the same as those delivered by brute force. Furthermore this is achieved consistently without the need for complex parameter tuning.

We contrast this to traditional blocking and LSH based approaches. The major drawback of their use is that whilst they can produce extremely good results, they can also produce extremely poor results. It was our observations of low recall given by these approaches that originally led us to experiment with M-trees. Without extensive ground truth or large amounts of clerking effort by human domain experts it is impossible to determine the quality of the links produced.

>>>>

\begin{figure}[!t]
  \centering
  \includegraphics[width=1.0\textwidth]{figures/linkage-process}
  \caption{Overview of the steps of the traditional record linkage
           process (left side) and our proposed metric space
           indexing based approach (right side), as described in
           Sect.~\ref{sec-intro}, where records from two datasets,
           $\mathbf{D}_A$ and $\mathbf{D}_B$, are being linked.}
           \label{fig-rl-process}
\end{figure}

@@

Metric space indexing (MSI) techniques provide indexing structures to
support the comparison of records in one set with those in another,
typically also offering similarity search operations. To create an MSI
data structure, it is necessary to define a distance measure between
records, with certain properties including the \emph{triangle
inequality}. Similarity search operations include
\emph{range-search(\textbf{q},d)}, where all records $\mathbf{r}$
in a database $\mathbf{D}$ within a distance $d$ of a query record
$\mathbf{q}$ are identified; \emph{nearest-neighbour(\textbf{q})},
returning the record $\mathbf{r} \in \mathbf{D}$ with smallest
distance to $\mathbf{q}$; and \emph{nearest-n(\textbf{q},n)}, returning
the $n$ closest records in $\mathbf{D}$ to $\mathbf{q}$.
\cite{Zezula2010} gives an overview of MSI techniques. Here we choose
one metric space indexing structure, the M-tree~\cite{paolociaccia2m},
and investigate its efficacy for record linkage.

@@

We address the following general record linkage problem: for two
datasets $\mathbf{D}_A$ and $\mathbf{D}_B$, we wish to find, for each
record in $\mathbf{D}_A$, all the records in $\mathbf{D}_B$ that match
it. We compare several algorithms for record linkage: traditional
blocking, one incomplete similarity search method, LSH-MinHash, and
one complete method, M-tree~\cite{paolociaccia2m}. We also use a
simple complete brute force technique as a baseline for comparison,
although this can only feasibly be applied to the smallest of our
datasets. All experiments have a number of parameters to configure
the search space and algorithm behaviour, including a
\emph{distance threshold}, $d$, specifying a maximum distance,
equivalent to a minimum similarity, for two records to be classified
as a link (i.e.\ referring to the same entity).

\smallskip
\textbf{Brute force:}
Two nested loops are used to compare every record in
$\mathbf{D}_A$ with every record in $\mathbf{D}_B$. Each pair is
classified as a link if the distance $d$ between the records is less
than the threshold. This approach is guaranteed to identify all links,
with complexity $ O(|\mathbf{D}_A| \times |\mathbf{D}_B|) $.

\smallskip
\textbf{Traditional blocking:}
The parameters of traditional blocking are the set of blocking keys
and the (optionally) phonetic
% Peter: normalisation?
encodings applied to each attribute. These are selected as described
in~\cite{Chr12b}, exploiting knowledge of the domain and of the
records being linked, and chosen with the intention of giving the
best possible results for the datasets. Each record in $\mathbf{D}_A$
is placed into the appropriate block based on its blocking key value.
The algorithm then iterates over the records in $\mathbf{D}_B$, and
for each one, compares it with each of the records from $\mathbf{D}_A$
in the block with the same blocking key value.

\smallskip
\textbf{LSH-MinHash:}
The parameters for LSH-Minhash are~\cite{Broder1997} \emph{shingle
size} ($l_{ss}$), \emph{band size} ($l_{bs}$) and \emph{number of
bands} ($l_{nb}$). Each record in $\mathbf{D}_A$ is placed in an
\emph{LSH-MinHash} data structure. First, all the attributes of the
record are concatenated, and the result \emph{shingled} into a set of
n-grams with $n = l_{ss}$. Next, a set of deterministically generated
hash functions are applied to each n-gram in the set and the smallest
result (the MinHash) of each hash application is added to a signature
for the record. The number of hashes used, and thus the size of the
signature, is set to $l_{nb} \times l_{bs}$. Finally, the signature is
split into $l_{nb}$ bands and the values from each band are hashed
again to create a number of keys. The original record is added to a
map associated with each of the keys.

To perform linkage, the algorithm iterates over the records in
$\mathbf{D}_B$. Each record is hashed using the procedure described
above, to obtain a set of keys. For each of these keys, the key is
looked up in the LSH-MinHash data structure, and the records from
$\mathbf{D}_A$ associated with it added to the result set. Finally,
the record from $\mathbf{D}_B$ is compared in turn with each record
in the result set using the same distance function and threshold, $d$,
as with the other approaches, with the pair being classified as a link
or non-link based on their distance.

In some circumstances, incomplete approaches such as traditional
blocking and LSH-MinHash can yield higher precision than complete
techniques. This can occur when the datasets contain a significant
number of non-matches that nonetheless have high similarity. In this
situation, the fact that an incomplete technique omits consideration
of some potential links can serve to improve precision, since a
classification decision based on a certain similarity threshold is
incorrect for high-similarity non-matches. By definition, recall can
never be higher for incomplete techniques.

%\smallskip
\textbf{M-tree:} Besides the distance threshold, $d$, the M-tree
linkage algorithm has no additional parameters. In a similar manner
to the \emph{LSH-MinHash} approach, each record in $\mathbf{D}_A$ is inserted into an M-tree. To perform linkage, the algorithm iterates
over each record in $\mathbf{b} \in \mathbf{D}_B$. A
\emph{range-search(\textbf{b},d)} operation is performed, passing the
distance threshold as the second parameter. All the returned records
are directly classified as links.

@@

\begin{table}[t]
\caption{Characteristics of datasets used in the experiments.}
 \label{table-datasets}
  \centering
  \begin{scriptsize}
  %\addtolength{\tabcolsep}{-0.5pt}
  \begin{tabular}{ccccc}
  \hline\noalign{\smallskip}
  Dataset~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities \\
    name(s)  & ~dataset $\mathbf{D}_A$~ & ~dataset $\mathbf{D}_B$~ &
    matching pairs & linked \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
  Cora & 1,295 & 1,295 & 17,184 & Publication--Publication\\
  ~Isle of Skye~ & 17,612 & 12,284& 2,900 & Birth--Death \\
  Kilmarnock & ~38,430~ & ~23,714~ & ~8,300~ & ~Birth--Death~ \\
  \noalign{\smallskip} \hline
  \end{tabular}
  \end{scriptsize}
\end{table}

@@

\smallskip
\textbf{Datasets:}
\label{sec-data}
We used three datasets from two domains in our experiments, as
summarised in Table~\ref{table-datasets}. The first is
\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer
to 112 machine learning publications. Cora is commonly used as a
benchmark dataset in the literature for assessing linkage algorithms.
Ground truth is provided via a unique \emph{paper\_id} identifier of
the form ``blum1993". In this experiment linkage is performed over
the same set of records.

The other two datasets are historical Scottish records of vital events
(birth, marriages and deaths) one registered on the \emph{Isle of
Skye}, a rural district, and the other records from
\emph{Kilmarnock}, an industrial town. These datasets were created by
historical demographers who extensively curated and linked both
datasets~\cite{reid2002,reid2006}. Both datasets include the names,
gender, addresses of individuals and their parents. Ground truth was
generated by the demographers based on their extensive domain
knowledge.

In all of our experiments we use a single distance metric which is the
sum of the Levenshtein~\cite{Levenshtein66} distances between the set
of attributes being compared.

@@

In this section, we perform linkage on the Cora dataset using all four
approaches presented in this paper: brute force, traditional blocking,
LSH and M-tree, using several selected configurations for the blocking
and LSH approaches. The distance threshold is varied between $0 \le d \le 250$~\footnote{Relatively high Levenshtein distances are included
since Cora includes a number of low-similarity true matches.}. For
traditional blocking, the following Cora attributes are used
individually as blocking keys: \emph{author}, \emph{title},
\emph{venue}, \emph{location}, \emph{publisher} and \emph{year}. We
also use a combined blocking key comprising all attributes.

\Cref{cora-quality} shows the precision, recall, and F-measure~\cite{Chr12} for various thresholds. As expected, low 
thresholds give high precision and low recall, and the reverse for high
threshold values. Brute force and M-tree give identical results, as
expected. The best linkage quality, with an F-measure of around 0.7,
is achieved by a number of linkers, including brute force, M-tree, blocking on \emph{authors} and blocking and on all attributes, and two
of the LSH configurations. All of these give similar overall results,
apart from blocking on \emph{authors}, which gives much better quality
at very high distance threshold values. This is due to the incomplete
nature of the approach, avoiding comparisons of significant numbers
of high-similarity non-matches and thus avoiding these becoming false
positives and keeping precision high.

For a more detailed investigation of selected linkers, the brute force
approach is used to establish a good threshold value for the Cora
dataset. The maximum F-measure is observed at a threshold value of
$d=70$. This value is dataset-dependent; for different datasets the
maximum F-measure will occur at different thresholds. In the rest of
this section we fix the threshold value at $d=70$.

\Cref{comparison-of-results-cora} shows greater detail for selected
linkers, showing the parameters for the experiment, the number of
distance comparisons made, the precision, recall and the F-measure
achieved by each algorithm. In the \emph{Linker} column the algorithm
name is followed by its parameters: for LSH the number of the bands
followed by the band size, and for traditional blocking the attributes
used for blocking. The number of distance comparisons is reported as a
machine-independent proxy for execution cost, since code profiling
shows that distance calculations are dominant.

@@

\begin{table}[ht]
\caption{Linkage quality on Cora dataset with distance threshold
  $d= 70$.}
  \label{comparison-of-results-cora}
  \centering
  \begin{footnotesize}
  \begin{tabular}{lrrrr} \hline\noalign{\smallskip}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68  \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
LSH-5-2            &  342,849	& 0.91	& 0.55	& 0.69       \\
LSH-10-2           &  513,947	& 0.88	& 0.57	& 0.69       \\
LSH-2-5            &  14,329	& 0.99	& 0.28	& 0.43      \\
LSH-5-5            &  22,057	& 0.99	& 0.36	& 0.53       \\
LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
LSH-2-10           &  4,711	& 1	& 0.15	& 0.27    \\
LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\
\noalign{\smallskip} \hline \noalign{\smallskip}
Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
Block-authors      &   11,039                   & 0.94      & 0.16   & 0.28       \\
Block-title        &   27,407                   & 0.95      & 0.42   & 0.58       \\
Block-venue        &   36,647                   & 0.85      & 0.29   & 0.44       \\
Block-location     & 1,009,957                 & 0.83      & 0.43   & 0.57       \\
Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

Birth records were linked to death records, separately for the Skye and
Kilmarnock datasets, using M-tree and a range of LSH configurations. It
was not computationally feasible to run the brute force linker. Of the
incomplete linkers, LSH was selected as it gave slightly better results
for Cora. The shingle size was set to 2 for all the LSH experiments
reported, as this was found to give good results and LSH was not
especially sensitive to this parameter. Results for other shingle sizes
are omitted from this paper for brevity. A lower range of distance
thresholds was explored, based on domain knowledge of the datasets.

\Cref{demography-quality} plots (a) and (b) show the M-tree precision,
recall, and F-measure for various thresholds. In both datasets, the best
F-measure values are obtained with a low distance threshold of 2. Plots
(c) and (d) compare the F-measure curves for M-tree with those obtained
from a range of LSH configurations. The best F-measure value for M-tree
is higher than that of any of the LSH configurations, for both datasets.
This demonstrates both the competitiveness of M-tree with respect to
linkage quality, and its important characteristic of being
parameter-free---the linkage quality is obtained without the need to
tuning for the dataset.

\Cref{comparison-of-results-demography-skye} and
\Cref{comparison-of-results-demography-kili} show greater detail for
selected linkers. In both cases the F-measure achieved is better for
M-tree than any of the LSH linkers. The better linkage quality achieved
by M-tree is largely due to recall for M-tree being much higher than for
any of the LSH configurations. In most cases, LSH out-performs M-tree in
terms of precision. More significantly, LSH linkage quality is heavily
dependent on the configuration parameters. For plausible settings for
the number of bands and band size, F-measure varies from 0.01 (extremely
poor) to 0.47 (relatively good) for Skye and from 0.03 to 0.49 for
Kilmarnock. In both cases \emph{LSH-10-2} performs best, but since this
is data-dependent there is no guarantee that these parameters would work
well with another dataset.

The number of distance comparisons varies dramatically among the various
linkers. M-tree always performs the most comparisons, since they are
intrinsic to the \emph{range-search} algorithm. The core part of the LSH
linker performs Jaccard similarity comparisons and hashing; distance
comparisons are only performed in the final step to determine whether a
candidate pair is a link. The LSH configurations yielding the best
results perform distance comparisons of the same order of magnitude as
M-tree. This indicates that the good LSH linkers return many candidates
beyond the distance threshold. Thus, in order to get good results, LSH
tends towards a brute force search over the candidate results. Despite
this, LSH is much faster due to the efficiency of the hashing process.

@@

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Skye dataset\label{skye-quality-mtree}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
\vspace{-6mm}
\caption{M-tree linkage on Kilmarnock dataset\label{kilmarnock-quality-mtree}}
\end{subfigure} \vspace{5mm}

\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
\vspace{-6mm}
\caption{F1 Measure on Skye dataset \\ with M-tree and all LSH configurations}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
\vspace{-6mm}
\caption{F1 Measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
\end{subfigure}

@@

\begin{table}[ht]
\caption{Linkage quality on Skye dataset with distance threshold
$d= 2$.}
\label{comparison-of-results-demography-skye}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrr} \hline\noalign{\smallskip}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\ \hline
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2  & 3,109,250                   & 0.63      & 0.03   & 0.06       \\

@@

LSH-10-5 & 129,873                    & 0.72      & 0.01   & 0.02       \\
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

\begin{table}[ht]
\caption{Linkage quality on Kilmarnock dataset with distance
threshold $d= 2$.}
\label{comparison-of-results-demography-kili}
\centering
\begin{footnotesize}
\begin{tabular}{lrrrrrr} \hline\noalign{\smallskip}
\parbox{0.2\textwidth}{\centering Linker} &
\parbox{0.2\textwidth}{\centering Comparisons} &
\parbox{0.17\textwidth}{\centering Precision} &
\parbox{0.17\textwidth}{\centering Recall} &
\parbox{0.17\textwidth}{\centering F1 Measure} \\
\noalign{\smallskip} \hline \noalign{\smallskip}
M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\
\noalign{\smallskip} \hline \noalign{\smallskip}
LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338                   & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848                   & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ 
\noalign{\smallskip} \hline
\end{tabular}
\end{footnotesize}
\end{table}

@@

In this paper we have demonstrated the efficacy of MSI in achieving
complete and efficient record linkage, without the need for complex
parameter tuning. In conclusion, this claim deserves some careful
unpacking. It is always possible to achieve high quality linkage using a
brute force approach. However the quadratic complexity of this approach
prevents its practical application for datasets of even moderate size.
We have shown that MSI techniques such as M-tree can deliver high
precision, high recall results that are the same as those delivered by
brute force. Furthermore this is achieved with fewer distance
comparisons, and consistently without the need for complex parameter
tuning.

We contrast this to traditional blocking and LSH-based approaches. Their
major drawback is that whilst they can produce extremely good results,
they can also produce extremely poor results. It was our observations of
low recall given by these approaches that originally led us to
experiment with M-trees.
